biological data preperation

### 2018-10-04 src/biological_data_prep.Rmd
short update to coordinates

The dataset is global and needs to be trimmed to contain just the area of interest. 
Plot all the occurance points
```{r map all occurrence points}
library(rworldmap)
map <- getMap(resolution = "low") #creates an object called map at low resoultion
plot(map, xlim = c(-180, -44), ylim =c(0,80), asp = 1) #the x and y lim are the long-lat bounds of the map
points(tidydata$decimalLongitude, tidydata$decimalLatitude) #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
```
now extract just the Atlantic Canada data.
The lon-lat study area (which captures the whole Atlantic Canadian EEZ up to the north of Resolution Island in the Hudson Strait/ Davis Strait) is -71.0,-43.0,62.0,38.0 
```{r map only atlantic canadian occurrences}
tidydata <- tidydata[tidydata$decimalLongitude > -71, ]
tidydata <- tidydata[tidydata$decimalLongitude < -43, ]
tidydata <- tidydata[tidydata$decimalLatitude > 38, ]
plot(map, xlim = c(-180, -44), ylim =c(0,80), asp = 1) 
points(tidydata$decimalLongitude, tidydata$decimalLatitude)
```
ok looks good...save what you have so far to a .csv
```{r save tidydata to csv}
write.csv(tidydata, file = "../data/bio/capelin_obis_tidy.csv", row.names = FALSE)
```

@@

```{r}
map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
plot(map2, xlim = c(-71, -43), ylim =c(38, 62), asp = 1, main = "All Occurrences", col = "cornsilk") #the x and y lim are the long-lat bounds of the map
points(tidydata$decimalLongitude, tidydata$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
dev.copy(png, "../output/bio/all_occurrences.png") #this prints a png of the plot
dev.off() #this turns off the print command
```
Note there is one point up by iceland that you should get rid of (Icelandic population thought to be seperate from Labrador, but it is unclear if this is true or not).

Map the institutioncode == ARC only data...
```{r map obs by institutuion}
ARC_obs <- tidydata[tidydata$institutioncode == "ARC", ]
map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
plot(map2, xlim = c(-71, -43), ylim =c(39, 62), asp = 1, main = "Arc Occurrences", col = "cornsilk") #the x and y lim are the long-lat bounds of the map
points(ARC_obs$decimalLongitude, ARC_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
dev.copy(png, "../output/bio/ARC_occurrences_all.png") #this prints a png of the plot
dev.off() #this turns off the print command

CAFF_obs <- tidydata[tidydata$institutioncode == "CAFF", ]
map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
plot(map2, xlim = c(-71, -43), ylim =c(39, 62), asp = 1, main = "CAFF Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
points(CAFF_obs$decimalLongitude, CAFF_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
dev.copy(png, "../output/bio/CAFF_occurrences_all.png") #this prints a png of the plot
dev.off() #this turns off the print command

DFOCENARC_obs <- tidydata[tidydata$institutioncode == "DFOCENARC", ]
map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
plot(map2, xlim = c(-71, -43), ylim =c(39, 62), asp = 1, main = "DFO Central & Arctic Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
points(DFOCENARC_obs$decimalLongitude, DFOCENARC_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
dev.copy(png, "../output/bio/DFOCENARC_occurrences_all.png") #this prints a png of the plot
dev.off() #this turns off the print command

DFOGulf_obs <- tidydata[tidydata$institutioncode == "DFOGulf", ]
map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
plot(map2, xlim = c(-71, -43), ylim =c(39, 62), asp = 1, main = "DFO Gulf Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
points(DFOGulf_obs$decimalLongitude, DFOGulf_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
dev.copy(png, "../output/bio/DFOGulf_occurrences_all.png") #this prints a png of the plot
dev.off() #this turns off the print command

DFOISDM_obs <- tidydata[tidydata$institutioncode == "DFOISDM", ]
map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
plot(map2, xlim = c(-71, -43), ylim =c(39, 62), asp = 1, main = "DFO ISDM Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
points(DFOISDM_obs$decimalLongitude, DFOISDM_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
dev.copy(png, "../output/bio/DFOISDM_occurrences_all.png") #this prints a png of the plot
dev.off() #this turns off the print command

DFOMTMS_obs <- tidydata[tidydata$institutioncode == "DFOMTMS", ]
map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
plot(map2, xlim = c(-71, -43), ylim =c(39, 62), asp = 1, main = "DFO Maritimes Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
points(DFOMTMS_obs$decimalLongitude, DFOMTMS_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
dev.copy(png, "../output/bio/DFOMTMS_occurrences_all.png") #this prints a png of the plot
dev.off() #this turns off the print command

DFONL_obs <- tidydata[tidydata$institutioncode == "DFONL", ]
map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
plot(map2, xlim = c(-71, -43), ylim =c(39, 62), asp = 1, main = "DFO Newfouundland & Labrador Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
points(DFONL_obs$decimalLongitude, DFONL_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
dev.copy(png, "../output/bio/DFONL_occurrences_all.png") #this prints a png of the plot
dev.off() #this turns off the print command

DFOQC_obs <- tidydata[tidydata$institutioncode == "DFOQC", ]
map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
plot(map2, xlim = c(-71, -43), ylim =c(39, 62), asp = 1, main = "DFO Quebec Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
points(DFOQC_obs$decimalLongitude, DFOQC_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
dev.copy(png, "../output/bio/DFOQC_occurrences_all.png") #this prints a png of the plot
dev.off() #this turns off the print command

ICES_obs <- tidydata[tidydata$institutioncode == "ICES", ]
map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
plot(map2, xlim = c(-71, -43), ylim =c(39, 62), asp = 1, main = "ICES Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
points(ICES_obs$decimalLongitude, ICES_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
dev.copy(png, "../output/bio/ICES_occurrences_all.png") #this prints a png of the plot
dev.off() #this turns off the print command

IMR_obs <- tidydata[tidydata$institutioncode == "IMR", ]
map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
plot(map2, xlim = c(-71, -43), ylim =c(39, 62), asp = 1, main = "IMR Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
points(IMR_obs$decimalLongitude, IMR_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
dev.copy(png, "../output/bio/IMR_occurrences_all.png") #this prints a png of the plot
dev.off() #this turns off the print command

MaineDMR_obs <- tidydata[tidydata$institutioncode == "MaineDMR", ]
map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
plot(map2, xlim = c(-71, -43), ylim =c(39, 62), asp = 1, main = "Maine DMR Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
points(MaineDMR_obs$decimalLongitude, MaineDMR_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
dev.copy(png, "../output/bio/MaineDMR_occurrences_all.png") #this prints a png of the plot
dev.off() #this turns off the print command

NAFO_obs <- tidydata[tidydata$institutioncode == "NAFO", ]
map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
plot(map2, xlim = c(-71, -43), ylim =c(39, 62), asp = 1, main = "NAFO Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
points(NAFO_obs$decimalLongitude, NAFO_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
dev.copy(png, "../output/bio/NAFO_occurrences_all.png") #this prints a png of the plot
dev.off() #this turns off the print command

NEFSC_obs <- tidydata[tidydata$institutioncode == "NEFSC", ]
map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
plot(map2, xlim = c(-71, -43), ylim =c(39, 62), asp = 1, main = "NEFSC Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
points(NEFSC_obs$decimalLongitude, NEFSC_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
dev.copy(png, "../output/bio/NEFSC_occurrences_all.png") #this prints a png of the plot
dev.off() #this turns off the print command

ROM_obs <- tidydata[tidydata$institutioncode == "ROM", ]
map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
plot(map2, xlim = c(-71, -43), ylim =c(39, 62), asp = 1, main = "Royal Ontario Museum Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
points(ROM_obs$decimalLongitude, ROM_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
dev.copy(png, "../output/bio/ROM_occurrences_all.png") #this prints a png of the plot
dev.off() #this turns off the print command

USNM_obs <- tidydata[tidydata$institutioncode == "USNM", ]
map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
plot(map2, xlim = c(-71, -43), ylim =c(39, 62), asp = 1, main = "National Museum of Natural History Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
points(USNM_obs$decimalLongitude, USNM_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
dev.copy(png, "../output/bio/USNM_occurrences_all.png") #this prints a png of the plot
dev.off() #this turns off the print command
```

@@

Now for  ROM there are more data points...
Use indentify function which lets you click on a point on the plot
```{r locate coordinates of iceland point}
x11() #this opens the plot i a window you can interact with
plot(map2, xlim = c(-71, -43), ylim =c(39, 62), asp = 1, main = "Royal Ontario Museum Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
points(ROM_obs$decimalLongitude, ROM_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude"))
identify(ROM_obs$decimalLongitude, ROM_obs$decimalLatitude, labels=ROM_obs$decimalLongitude) #decimalLongitude chosen as plot identifier as the values are unique
```
The point that needs to be removed is located at -44.96667. Note this is not the full value so use some other locators to remove the row (you could just used id, but pick multiple others just to learn)

```{r remove iceland point}
tidydata <- tidydata[!(tidydata$resname == "Fish specimens" & tidydata$datecollected == "1965-08-13 12:00:00+01"), ]
```

@@

```{r map occurrences less iceland}
plot(map2, xlim = c(-71, -43), ylim =c(38, 62), asp = 1, main = "All Occurrences without Iceland", col = "cornsilk") #the x and y lim are the long-lat bounds of the map
points(tidydata$decimalLongitude, tidydata$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
dev.copy(png, "../output/bio/all_occurrences_less_iceland.png") #this prints a png of the plot
dev.off() #this turns off the print command
```
great! write what you have so far to a csv
```{r}
write.csv(tidydata, file = "../data/bio/capelin_obis_tidy.csv", row.names = FALSE)
```

@@

Uh oh there is a lot. A bet a bunch of them are in the Hudson Strait (which is outside the NAFO Zone).Map the nafo_na points...
```{r}
plot(map2, xlim = c(-71, -43), ylim =c(38, 62), asp = 1, main = "Occurences missing a NAFO Zone", col = "cornsilk") #the x and y lim are the long-lat bounds of the map
points(nafo_na$decimalLongitude, nafo_na$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
```
ok yes - Hudson Strait has many, but there are some near the coast that are probably just missed out due to minor differences in the point location/shapefile boundaries.
Start with the Hudson Strait ones as this is easy - anything above 56N, nafo_zone == "HudsonStrait"

```{r}
adddata$nafo_zone <- as.character(adddata$nafo_zone)
adddata[is.na(adddata)] <- "xx"
adddata$nafo_zone[adddata$nafo_zone == "xx" & adddata$decimalLatitude > 56] <- "HudsonStrait"
```
and map to check
```{r}
nafo_na <- subset(adddata, nafo_zone == "xx")
plot(map2, xlim = c(-71, -43), ylim =c(38, 62), asp = 1, main = "Occurences missing a NAFO Zone", col = "cornsilk") #the x and y lim are the long-lat bounds of the map
points(nafo_na$decimalLongitude, nafo_na$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
```
so now i need to account for the others (either as occuring on land or just in an area that is missing a NAFO Zone tag) which is more tricky... i might just do this bit in ArcGIS quickly.
```{r}
write.csv(adddata, file = "../data/bio/capelin_additional_tidy.csv", row.names = FALSE)
```
So in ArcGIS I saw that these points fall on land. Some are extremely close to the coast, some further inland. There are only 132 of thesepoints, so I will just delete them from the database.
```{r}
adddata$nafo_zone[adddata$nafo_zone == "xx"] <- NA #make all xx values NA
adddata <- adddata[!is.na(adddata$nafo_zone), ]
write.csv(adddata, file = "../data/bio/capelin_additional_tidy.csv", row.names = FALSE)
```

@@

The warm/cold year index will be taken from a [DFO CSAS report](http://www.dfo-mpo.gc.ca/csas-sccs/Publications/SAR-AS/2018/2018_039-eng.pdf)

>>>>

The dataset is global and needs to be trimmed to contain just the area of interest. 
Plot all the occurance points
```{r map all occurrence points}
library(rworldmap)
map <- getMap(resolution = "low") #creates an object called map at low resoultion
plot(map, xlim = c(-180, -44), ylim =c(0,80), asp = 1) #the x and y lim are the long-lat bounds of the map
points(tidydata$decimalLongitude, tidydata$decimalLatitude) #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
```
now extract just the Atlantic Canada data.
The lon-lat study area (which captures the whole Atlantic Canadian EEZ up to the north of Resolution Island in the Hudson Strait/ Davis Strait) is -70.0,-43.0,70.0,38.0 
```{r map only atlantic canadian occurrences}
tidydata <- tidydata[tidydata$decimalLongitude > -70, ]
tidydata <- tidydata[tidydata$decimalLongitude < -43, ]
tidydata <- tidydata[tidydata$decimalLatitude > 38, ]
plot(map, xlim = c(-180, -42), ylim =c(0,80), asp = 1) 
points(tidydata$decimalLongitude, tidydata$decimalLatitude)
```
ok looks good...save what you have so far to a .csv
```{r save tidydata to csv}
write.csv(tidydata, file = "../data/bio/capelin_obis_tidy.csv", row.names = FALSE)
```

@@

```{r}
map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
plot(map2, xlim = c(-70, -43), ylim =c(38, 70), asp = 1, main = "All Occurrences", col = "cornsilk") #the x and y lim are the long-lat bounds of the map
points(tidydata$decimalLongitude, tidydata$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
dev.copy(png, "../output/bio/all_occurrences.png") #this prints a png of the plot
dev.off() #this turns off the print command
```
Note there is one point up by iceland that you should get rid of (Icelandic population thought to be seperate from Labrador, but it is unclear if this is true or not).

Map the institutioncode == ARC only data...
```{r map obs by institutuion}
ARC_obs <- tidydata[tidydata$institutioncode == "ARC", ]
map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
plot(map2, xlim = c(-70, -43), ylim =c(39, 70), asp = 1, main = "Arc Occurrences", col = "cornsilk") #the x and y lim are the long-lat bounds of the map
points(ARC_obs$decimalLongitude, ARC_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
dev.copy(png, "../output/bio/ARC_occurrences_all.png") #this prints a png of the plot
dev.off() #this turns off the print command

CAFF_obs <- tidydata[tidydata$institutioncode == "CAFF", ]
map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
plot(map2, xlim = c(-70, -43), ylim =c(39, 70), asp = 1, main = "CAFF Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
points(CAFF_obs$decimalLongitude, CAFF_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
dev.copy(png, "../output/bio/CAFF_occurrences_all.png") #this prints a png of the plot
dev.off() #this turns off the print command

DFOCENARC_obs <- tidydata[tidydata$institutioncode == "DFOCENARC", ]
map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
plot(map2, xlim = c(-70, -43), ylim =c(39, 70), asp = 1, main = "DFO Central & Arctic Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
points(DFOCENARC_obs$decimalLongitude, DFOCENARC_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
dev.copy(png, "../output/bio/DFOCENARC_occurrences_all.png") #this prints a png of the plot
dev.off() #this turns off the print command

DFOGulf_obs <- tidydata[tidydata$institutioncode == "DFOGulf", ]
map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
plot(map2, xlim = c(-70, -43), ylim =c(39, 70), asp = 1, main = "DFO Gulf Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
points(DFOGulf_obs$decimalLongitude, DFOGulf_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
dev.copy(png, "../output/bio/DFOGulf_occurrences_all.png") #this prints a png of the plot
dev.off() #this turns off the print command

DFOISDM_obs <- tidydata[tidydata$institutioncode == "DFOISDM", ]
map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
plot(map2, xlim = c(-70, -43), ylim =c(39, 70), asp = 1, main = "DFO ISDM Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
points(DFOISDM_obs$decimalLongitude, DFOISDM_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
dev.copy(png, "../output/bio/DFOISDM_occurrences_all.png") #this prints a png of the plot
dev.off() #this turns off the print command

DFOMTMS_obs <- tidydata[tidydata$institutioncode == "DFOMTMS", ]
map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
plot(map2, xlim = c(-70, -43), ylim =c(39, 70), asp = 1, main = "DFO Maritimes Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
points(DFOMTMS_obs$decimalLongitude, DFOMTMS_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
dev.copy(png, "../output/bio/DFOMTMS_occurrences_all.png") #this prints a png of the plot
dev.off() #this turns off the print command

DFONL_obs <- tidydata[tidydata$institutioncode == "DFONL", ]
map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
plot(map2, xlim = c(-70, -43), ylim =c(39, 70), asp = 1, main = "DFO Newfouundland & Labrador Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
points(DFONL_obs$decimalLongitude, DFONL_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
dev.copy(png, "../output/bio/DFONL_occurrences_all.png") #this prints a png of the plot
dev.off() #this turns off the print command

DFOQC_obs <- tidydata[tidydata$institutioncode == "DFOQC", ]
map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
plot(map2, xlim = c(-70, -43), ylim =c(39, 70), asp = 1, main = "DFO Quebec Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
points(DFOQC_obs$decimalLongitude, DFOQC_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
dev.copy(png, "../output/bio/DFOQC_occurrences_all.png") #this prints a png of the plot
dev.off() #this turns off the print command

ICES_obs <- tidydata[tidydata$institutioncode == "ICES", ]
map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
plot(map2, xlim = c(-70, -43), ylim =c(39, 70), asp = 1, main = "ICES Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
points(ICES_obs$decimalLongitude, ICES_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
dev.copy(png, "../output/bio/ICES_occurrences_all.png") #this prints a png of the plot
dev.off() #this turns off the print command

IMR_obs <- tidydata[tidydata$institutioncode == "IMR", ]
map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
plot(map2, xlim = c(-70, -43), ylim =c(39, 70), asp = 1, main = "IMR Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
points(IMR_obs$decimalLongitude, IMR_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
dev.copy(png, "../output/bio/IMR_occurrences_all.png") #this prints a png of the plot
dev.off() #this turns off the print command

MaineDMR_obs <- tidydata[tidydata$institutioncode == "MaineDMR", ]
map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
plot(map2, xlim = c(-70, -43), ylim =c(39, 70), asp = 1, main = "Maine DMR Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
points(MaineDMR_obs$decimalLongitude, MaineDMR_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
dev.copy(png, "../output/bio/MaineDMR_occurrences_all.png") #this prints a png of the plot
dev.off() #this turns off the print command

NAFO_obs <- tidydata[tidydata$institutioncode == "NAFO", ]
map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
plot(map2, xlim = c(-70, -43), ylim =c(39, 70), asp = 1, main = "NAFO Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
points(NAFO_obs$decimalLongitude, NAFO_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
dev.copy(png, "../output/bio/NAFO_occurrences_all.png") #this prints a png of the plot
dev.off() #this turns off the print command

NEFSC_obs <- tidydata[tidydata$institutioncode == "NEFSC", ]
map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
plot(map2, xlim = c(-70, -43), ylim =c(39, 70), asp = 1, main = "NEFSC Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
points(NEFSC_obs$decimalLongitude, NEFSC_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
dev.copy(png, "../output/bio/NEFSC_occurrences_all.png") #this prints a png of the plot
dev.off() #this turns off the print command

ROM_obs <- tidydata[tidydata$institutioncode == "ROM", ]
map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
plot(map2, xlim = c(-70, -43), ylim =c(39, 70), asp = 1, main = "Royal Ontario Museum Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
points(ROM_obs$decimalLongitude, ROM_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
dev.copy(png, "../output/bio/ROM_occurrences_all.png") #this prints a png of the plot
dev.off() #this turns off the print command

USNM_obs <- tidydata[tidydata$institutioncode == "USNM", ]
map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
plot(map2, xlim = c(-70, -43), ylim =c(39, 70), asp = 1, main = "National Museum of Natural History Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
points(USNM_obs$decimalLongitude, USNM_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
dev.copy(png, "../output/bio/USNM_occurrences_all.png") #this prints a png of the plot
dev.off() #this turns off the print command
```

@@

Now for  ROM there are more data points...
Use indentify function which lets you click on a point on the plot
```{r locate coordinates of iceland point}
x11() #this opens the plot i a window you can interact with
plot(map2, xlim = c(-70, -43), ylim =c(38, 70), asp = 1, main = "Royal Ontario Museum Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
points(ROM_obs$decimalLongitude, ROM_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude"))
identify(ROM_obs$decimalLongitude, ROM_obs$decimalLatitude, labels=ROM_obs$decimalLongitude) #decimalLongitude chosen as plot identifier as the values are unique
```
The point that needs to be removed is located at -44.96667 and -53,483333. Note this is not the full value so use some other locators to remove the row (you could just used id, but pick multiple others just to learn)

```{r remove iceland point}
tidydata <- tidydata[!(tidydata$resname == "Fish specimens" & tidydata$datecollected == "1965-08-13 12:00:00+01"), ]
tidydata <- tidydata[!(tidydata$resname == "Fish specimens" & tidydata$datecollected == "1967-10-25 12:00:00+01"), ]
tidydata <- tidydata[!(tidydata$resname == "Ichthyology Collection - Royal Ontario Museum" & tidydata$datecollected == "1967-10-25 12:00:00+01"), ]
```

@@

```{r map occurrences less iceland}
plot(map2, xlim = c(-70, -43), ylim =c(38, 70), asp = 1, main = "All Occurrences without Iceland", col = "cornsilk") #the x and y lim are the long-lat bounds of the map
points(tidydata$decimalLongitude, tidydata$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
dev.copy(png, "../output/bio/all_occurrences_less_iceland.png") #this prints a png of the plot
dev.off() #this turns off the print command
```
great! write what you have so far to a csv
```{r}
write.csv(tidydata, file = "../data/bio/capelin_obis_tidy.csv", row.names = FALSE)
```

@@

Uh oh there is a lot. A bet a bunch of them are in the Hudson Strait (which is outside the NAFO Zone).Map the nafo_na points...
```{r}
plot(map2, xlim = c(-70, -43), ylim =c(38, 70), asp = 1, main = "Occurences missing a NAFO Zone", col = "cornsilk") #the x and y lim are the long-lat bounds of the map
points(nafo_na$decimalLongitude, nafo_na$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
```
ok yes - Hudson Strait has many, but there are some near the coast that are probably just missed out due to minor differences in the point location/shapefile boundaries.
Start with the Hudson Strait ones as this is easy - anything above 56N, nafo_zone == "HudsonStrait"

```{r}
adddata$nafo_zone <- as.character(adddata$nafo_zone)
adddata[is.na(adddata)] <- "xx"
adddata$nafo_zone[adddata$nafo_zone == "xx" & adddata$decimalLatitude > 56] <- "HudsonStrait"
```
and map to check
```{r}
nafo_na <- subset(adddata, nafo_zone == "xx")
plot(map2, xlim = c(-70, -43), ylim =c(38, 70), asp = 1, main = "Occurences missing a NAFO Zone", col = "cornsilk") #the x and y lim are the long-lat bounds of the map
points(nafo_na$decimalLongitude, nafo_na$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
```
so now i need to account for the others (either as occuring on land or just in an area that is missing a NAFO Zone tag) which is more tricky... i might just do this bit in ArcGIS quickly.
```{r}
write.csv(adddata, file = "../data/bio/capelin_additional_tidy.csv", row.names = FALSE) #just write all the data you have
```

So in ArcGIS I saw that some points are extremely close to the coast, but all lie in 4T.
```{r}
adddata$nafo_zone[adddata$nafo_zone == "xx"] <- "4T" #make all xx values 4T
nafo_na <- subset(adddata, nafo_zone == "xx")
plot(map2, xlim = c(-70, -43), ylim =c(38, 70), asp = 1, main = "Occurences missing a NAFO Zone", col = "cornsilk") #the x and y lim are the long-lat bounds of the map
points(nafo_na$decimalLongitude, nafo_na$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
```
all sorted!

```{r}
write.csv(adddata, file = "../data/bio/capelin_additional_tidy.csv", row.names = FALSE)
```

@@

The warm/cold year index will be taken from a [DFO CSAS report](http://www.dfo-mpo.gc.ca/csas-sccs/Publications/SAR-AS/2018/2018_039-eng.pdf) - Fig 8
